---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(knitr)
```

# walkboutr

<!-- badges: start -->
<!-- badges: end -->

The goal of walkboutr is to process GPS and accelerometry data into walking bouts. walkboutr will either return the original dataset along with bout labels and categories, or a summarized, de-identified dataset that can be shared for collaboration. 

## Installation

You can install the development version of walkboutr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("rwalkbout/walkboutr")
```

## Example

This is an example of simulated data that could be processed by walkboutr. The GPS data contain the required columns: time, latitude, longitude, speed. The accelerometry data contain the required columns: time, accerometry counts.
These data have no extra columns, do not contain NAs, and don't have negative speeds or accelerometry counts. All times are also in date-time format. 


```{r simulated data}
library(walkboutr)
gps_data <- generate_walking_in_seattle_gps_data() # this will generate sample GPS data
accelerometry_counts <- make_full_day_bout_without_metadata() # this will generate sample accelerometry data 

(head(gps_data))
(head(accelerometry_counts))

```

Now that we have sample data, we can look at how the walkboutr package works. There are two top level functions that will allow us to generate either (1) a dataset with bouts and bout categories with all of our original data included, or (2) a summary dataset that is completely de-identified and shareable for research purposes. 

#### Walk bout dataset including original data

```{r example 1}
# walk_bouts <- identify_walk_bouts_in_gps_and_accelerometry_data(gps_data,accelerometry_counts)
# (head(walk_bouts))
```

#### Summarized walk bout dataset 
This dataset is a set of labelled bouts that are categorized (`bout_category`) and contains information on bout specific median speed (`median_speed`), the start time of the bout (`bout_start`), the duration of the bout (in minutes for computational ease, `duration`), and a flag for whether the bout came from a dataset with a complete day worth of data (`complete_day`). 
```{r example 2}
# summary <- summarize_walk_bouts(walk_bouts)
# (head(summary))
```








You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
