---
output: 
  github_document:
      toc: true
      toc_depth: 2
always_allow_html: true
fontsize: 14pt
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(knitr)
library(kableExtra)
```

# walkboutr

<!-- badges: start -->
<!-- badges: end -->

The goal of walkboutr is to process GPS and accelerometry data into walking bouts. walkboutr will either return the original dataset along with bout labels and categories, or a summarized, de-identified dataset that can be shared for collaboration. 

### Installation

You can install the development version of walkboutr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("rwalkbout/walkboutr")
```

### Basic Usage

##### Simulated sample data

This is an example of simulated data that could be processed by walkboutr. The GPS data contain the required columns: time, latitude, longitude, speed. The accelerometry data contain the required columns: time, accerometry counts.
These data have no extra columns, do not contain NAs, and don't have negative speeds or accelerometry counts. All times are also in date-time format. 

```{r simulated data}
library(walkboutr)
# generate sample gps data:
gps_data <- generate_walking_in_seattle_gps_data() 
# generate sample accelerometry data:
accelerometry_counts <- make_full_day_bout_without_metadata() 
```

GPS data: 
```{r simulated data: gps, echo = FALSE}
kable(head(gps_data), booktabs = TRUE)  %>% 
  kable_styling(latex_options = "striped", full_width = TRUE, font_size = 12, position = "center")
```

Accelerometry data:
```{r simulated data: acc, echo = FALSE}
kable(head(accelerometry_counts), booktabs = TRUE)  %>% 
  kable_styling(latex_options = "striped", full_width = TRUE, font_size = 12, position = "center")
```

<p>
<p>
Now that we have sample data, we can look at how the walkboutr package works. There are two top level functions that will allow us to generate either (1) a dataset with bouts and bout categories with all of our original data included, or (2) a summary dataset that is completely de-identified and shareable for research purposes. 


##### Walk bout dataset including original data

```{r example 1, message = FALSE, warning = FALSE, results = 'hide'}
walk_bouts <- identify_walk_bouts_in_gps_and_accelerometry_data(gps_data,accelerometry_counts)
```


```{r example printout, echo = FALSE}
kable(head(walk_bouts)) %>% 
  kable_styling(latex_options = "striped") %>% 
  kable_styling(font_size = 12) %>% 
  kable_styling(position = "center")
```


##### Summarized walk bout dataset 
This dataset is a set of labelled bouts that are categorized (`bout_category`) and contains information on bout specific median speed (`median_speed`), the start time of the bout (`bout_start`), the duration of the bout (in minutes for computational ease, `duration`), and a flag for whether the bout came from a dataset with a complete day worth of data (`complete_day`). 
```{r example 2, message = FALSE, warning = FALSE}
summary <- summarize_walk_bouts(walk_bouts)
```

```{r example 2 printout, echo = FALSE}
kable(head(summary)) %>% 
  kable_styling(latex_options = "striped", font_size = 12, position = "center")
summary <- summarize_walk_bouts(walk_bouts)
summary_1 <- summary %>% dplyr::filter(bout == 1)
```

In this example, we have `r nrow(summary)` bout(s), and each bout has a label. Bout 1 occurred on `r format(summary_1$bout_start, format = "%Y.%m.%d")` and has a complete day worth of data (`complete_day` = `r summary_1$complete_day`) and a start time of `r format(summary_1$bout_start, format = "%H:%M:%S")`. This bout lasted `r summary_1$duration` minutes, or `r summary_1$duration/60 %>% round(2)` hours. The bout is a non walk bout because the participant was moving too slowly for this walk to be considered walking. 

For more information on bout categories and how these are assigned, please see the vignette titled XXX. 


``` {r plot, include = FALSE}
bouts <- process_accelerometry_counts_into_bouts(accelerometry_counts)
gps_epochs <- process_gps_data_into_gps_epochs(gps_data)
all_bouts <- process_bouts_and_gps_epochs_into_walkbouts(bouts, gps_epochs)
```

